Let There Be Light: \\ Early Observations of the Normal Type Ia Supernova \abc\ \\Show No Evidence for a Dark Period

### 2017-08-03 doc/ms.tex
Update manuscript to reflect Nadia's comments

@@

Early observations of Type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia, \abc. The initial detection was made only $0.18\pm^{0.07}_{0.15}$~d after SN first light. In the $\sim$24 hr after discovery, \abc\ rose by $\sim$2 mag, following a near-linear rise in flux for $\sim$4 d. Assuming the photospheric velocity follows a $t^{-0.22}$ power law, the time of explosion for \abc\ is approximately equal to the time of first light. Strong \ion{C}{2} absorption is detected in the early spectra of \abc, before disappearing after $\sim$7~d. Unlike SN\,2011fe, the $(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after explosion. We show that the fast, near-linear rise in flux for \abc\ cannot be explained by either SN shock breakout and the associated, subsequent cooling or the collision of the SN ejecta with a stellar companion. Instead, we argue that the early characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors, and (iii) the strong lines from ionized carbon in the earliest spectra, are the result of either vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or the interaction of the ejecta with diffuse material, or a combination of the two. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.

@@

Using the Supernova Identification (SNID; \citealt{2007ApJ...666.1024B}) package, 
we find the low-resolution spectrum of \abc\ at $+18.8$ d is best matched by normal SNe Ia. Several characteristic features of a SN
Ia, such as \ion{Si}{2}, \ion{S}{2}, can be easily identified in the
spectra of \abc\ (Figure \ref{fig:spec_seq}).

@@

The early spectra of \abc\ exhibit unusually strong absorption due 
to \ion{C}{2} $\lambda\lambda$6580, 7234. We highlight the evolution 
of these spectral features in Figure~\ref{fig:carbon}. From these 
spectra we see that \ion{C}{2}\,$\lambda$6580 is as strong as 
\ion{Si}{2}\,$\lambda$6355 at $t \approx -16 \; \mathrm{d}$. The 
strength of the \ion{C}{2} lines declines with time to the point 
where they are no longer detectible more than 1 wk after explosion.

\begin{figure}[]
  \centering
  \includegraphics[width=0.49\textwidth]{CarbonFeature.pdf}
  \caption{Evolution of the \ion{C}{2} features observed in the 
  early spectra of \abc. The raw spectra are shown in orange, while 
  the solid blue lines show the best-fit models including gaussian 
  components for each line and a linear component for the continuum 
  (see text for further details). The dark grey vertical lines show 
  the measured line centers, indicating the decline in the 
  photosphere velocity in the $\sim$wk after explosion (note that 
  \ion{C}{2}\,$\lambda$7234 is not detected in the $-10.9$ d 
  spectrum). The phase of each spectrum relative to 
  $T_{B,\mathrm{max}}$ is shown to the right of each spectrum.
  }
  \label{fig:carbon}
\end{figure}

@@

\begin{itemize}
    \item Extrapolation of the early light curve shows that the initial detection of \abc\ occurred only 0.15 d after the time of first light.

@@

>>>>

\author{N.~Blagorodnova}
\affil{Division of Physics, Mathematics, and Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}

@@

Early observations of Type Ia supernovae (SNe) provide a unique probe of their progenitor systems and explosion physics. Here, we report the intermediate Palomar Transient Factory (iPTF) discovery of an extraordinarily young SN Ia, \abc. The initial detection was made only $0.15\pm^{0.07}_{0.15}$~d after SN first light. In the $\sim$24 hr after discovery, \abc\ rose by $\sim$2 mag, following a near-linear rise in flux for $\sim$4 d. Assuming the photospheric velocity follows a $t^{-0.22}$ power law, the time of explosion for \abc\ is approximately equal to the time of first light. Strong \ion{C}{2} absorption is detected in the early spectra of \abc, before disappearing after $\sim$7~d. Unlike SN\,2011fe, the $(B-V)_0$ colors of \abc\ are blue and nearly constant in the days after explosion. We show that the fast, near-linear rise in flux for \abc\ cannot be explained by either SN shock breakout and the associated, subsequent cooling or the collision of the SN ejecta with a stellar companion. Instead, we argue that the early characteristics of \abc, including: (i) the rapid, near-linear rise, (ii) the non-evolving blue colors, and (iii) the strong lines from ionized carbon in the earliest spectra, are the result of either vigorous mixing of radioactive $^{56}$Ni in the SN ejecta, or the interaction of the ejecta with diffuse material, or a combination of the two. In the next few years, dozens of very young \textit{normal} SNe Ia will be discovered, and observations similar to those presented here will constrain the white dwarf explosion mechanism.

@@

Using the Supernova Identification (\texttt{SNID}; \citealt{2007ApJ...666.1024B}) package, 
we find the low-resolution spectrum of \abc\ at $+18.8$ d is best matched by normal SNe Ia. Several characteristic features of a SN
Ia, such as \ion{Si}{2}, \ion{S}{2}, can be easily identified in the
spectra of \abc\ (Figure \ref{fig:spec_seq}).

@@

The early spectra of \abc\ exhibit unusually strong absorption due 
to \ion{C}{2} $\lambda\lambda$6580, 7234. We highlight the evolution 
of these spectral features in Figure~\ref{fig:carbon}. From these 
spectra we see that \ion{C}{2}\,$\lambda$6580 is as strong as 
\ion{Si}{2}\,$\lambda$6355 at $t \approx -16 \; \mathrm{d}$. The 
strength of the \ion{C}{2} lines declines with time, and by $t \approx -10 \; \mathrm{d}$ \ion{C}{2} in no longer detectable.

\begin{figure}[]
  \centering
  \includegraphics[width=0.49\textwidth]{CarbonFeature.pdf}
  \caption{Evolution of the \ion{C}{2} features observed in the 
  early spectra of \abc. The raw spectra are shown in orange, while 
  the solid blue lines show the best-fit models including gaussian 
  components for each line and a linear component for the continuum 
  (see text for further details). The dark grey vertical lines show 
  the measured line centers, indicating the decline in the 
  photosphere velocity in the $\sim$7~d after explosion (note that 
  \ion{C}{2}\,$\lambda$7234 is not detected in the $-10.9$ d 
  spectrum). The phase of each spectrum relative to 
  $T_{B,\mathrm{max}}$ is shown to the right of each spectrum.
  }
  \label{fig:carbon}
\end{figure}

@@

\begin{itemize}
    \item Extrapolation of the early light curve shows that the initial detection of \abc\ occurred only $0.15\pm^{0.07}_{0.15}$~d after the time of first light.

@@

\facility{DCT, Gemini:Gillett, Hale, Keck:I, Keck:II, LCOGT, PO:1.2\,m, PO:1.5\,m, VLT, Swift, OANSPM:HJT}

\software{\texttt{PTFIDE} \citep{Masci2017PASP}, \texttt{SALT2}, \texttt{SNID}, \texttt{UVOTSOURCE}}.
